Overlapping Batches for the Assessment of Solution Quality in Stochastic Programs

### 2014-12-22 omrp_lb_fl.tex
R2 minor edits, cont'd thru p. 19

In order to apply overlapping batches to MRP, we need to keep track of solutions to sampling problems (SP$_m$) for each batch of size $m$.  
Toward this end, let $B(i)$ denote the set of batches $j \in \left\{1, 2, \dots, \nb \right\}$ observation $\xiti$ is used in, $i = 1, 2, \dots, n$.  
See Figure \ref{fig:overlap_nonint} for an example with $n=18$, $m = 6$, and $\gamma = 2$.  
Here, the first batch uses observations $\xit^1, \xit^2, \xit^3, \xit^4, \xit^5, \xit^6$, the second batch uses $\xit^3, \xit^4, \xit^5, \xit^6, \xit^7, \xit^8$, and so on.  
So, $B(1) = \{1\}$, $B(3) = \{1,2\}$, and $B(7)=\{2,3,4\}$.  
As before, we use $\xs_j$ to denote an optimal solution to sampling problem (SP$_m$) formed using the $j$th batch. 
The sampling problem (SP$_m$) formed using the $j$th batch might have multiple optimal solutions. 
We use $\xs_j \in \argmin_{\x \in X} \frac{1}{m} \sum_{i=1}^m f(\x,\xit^{\gamma(j-1) + i})$, $j = 1, 2, \dots, \nb$ to denote one optimal solution.

@@

The optimality gap estimator for each batch $\gb_{lj}$ is defined like (\ref{eq:mrp_gap}) for general values of the nonoverlap parameter $\gamma_l$.  
We just removed the minimization in (\ref{eq:mrp_gap}) and used directly an optimal solution of $\xs_j$ of batch $j$.  
The overall mean, $\gbb_l$, is defined a little differently.  
Here, $\gbb_l$ still uses each observation $i = 1, 2, \dots, n_l$ but also makes use of {\it all} the information collected throughout the batches.  
That is, if observation $\xiti$ is used in $|B(i)|$ number of batches, all the optimal solutions $\xs_j$ corresponding to each batch $j \in B(i)$ are used for the lower bound estimator.  
Then, $VG_l$ is defined in a similar fashion for the overlapping batches variance estimator (\ref{eq:variance_gamma}).

The motivation behind forming $\gbb_l$ by taking into account $B(i)$ is to give each observation $\xit^i$ equal weight. 
This is because we want to be as close to a standard mean estimator (e.g., $\ybb=\frac{1}{n}\sum_{i=1}^{n} y^i$) as possible.
With this definition of $\gbb_l$, when $\xs_j= \xs$ for all $j= 1, 2, \dots, \nbl$, we have $\gbb_l$ equal to its nonoptimized counterpart given in (\ref{eq:dbb}), which is the standard mean estimator. 
This fact is key in establishing the theoretical results presented in the next section. 
This choice of $\gbb_l$, however, may not be guaranteed to be nonnegative.
In practice, $\max\{\gbb_l,0\}$ can be used as the point estimator of optimality gap to ensure nonnegativity, and $VG_l$ can be calculated the same way given in (\ref{eq:vg}), using $\gbb_l$, to obtain improved interval estimators via overlapping. 
%In our computational results, however, we did not observe any negative values of $\gbb_l$.

@@

In this section, we provide conditions under which several asymptotic properties hold for OMRP estimators. 
First, we show the consistency of the point estimators of optimality gap, $\gbb_l$, and variance, $n_l VG_l$. 
Then, we provide conditions under which the OMRP variance estimator achieves asymptotically lower variances relative to the MRP estimator.  
Finally, we show the asymptotic validity of the OMRP interval estimator.

@@

\begin{description}
	\item[A1] Samples of the random variable $\xit$ are i.i.d.
	\item[A2] $\sup_{\x \in X} \e{f(\x,\xit)^{4}} < \infty$.
	\item[A3] (\ref{eq:sto_prog}) has a unique optimal solution $\xs$ and $\p{\xs_m \neq \xs} \rightarrow 0$ exponentially fast as $m \rightarrow \infty$.  
           That is, there exists constants $M > 0$ and $c > 0$ such that $\log(\p{\xs_m \neq \xs}) \leq -cm$ for all $m > M$.
\end{description}

First, we restrict our attention to i.i.d.\ sampling by A1.
The results for overlapping batches from the simulation output analysis indicate that $\var{n_l VG_l}$ may be reduced by overlapping, thus requiring at least finite fourth moments. 
We will formally show this result in \S~\ref{ssec:var_reduct}.
Our proof relies on an exponential rate of convergence of the sampled solution to the true solution of the problem, specified by A3.  
Assumption A3 can be satisfied for different classes of problems with unique optimal solutions.  
\citet{shapiro2000rate} establish exponential rates of convergence for a class of (\ref{eq:sto_prog}), which include two-stage stochastic linear programs with recourse when the support of $\xit$ is finite.   
\citet{kleywegt2002sample} show that assumption A3 can be satisfied for discrete stochastic programs (i.e., when $X$ is a finite set) under the condition that the moment generating function of $f(\xs,\xit) - f(\x,\xit)$ for $\x \in X \backslash \{\xs\}$ is finite valued on $\R$.  
This class of problems includes stochastic integer programs with recourse.
  We present computational results on both stochastic linear and integer programs with recourse in \S \ref{sec:comp}.

% \begin{description}
% 	\item[A'1] (\ref{eq:sto_prog}) has a unique and sharp optimal solution, i.e., there is a constant $C$ such that $\e{f(\x,\xi)} \geq \e{f(\xs,\xi)} + C\Vert \x - \xs\Vert$ for all $\x \in X$
% 	\item[A'2] (\ref{eq:sto_prog}) is Lipschitz continuous almost surely, i.e., there is a constant $K$ such that $|f(\x_1,\xit) - f(\x_2,\xit)| \leq K \Vert \x_1 - \x_2 \Vert$ for all $\x_1,\x_2 \in X$.

% Some results require an additional assumption:
% \begin{description}
% 	\item[A4] (\ref{eq:sto_prog}) has a unique optimal solution.
% \end{description}

@@

Before we begin establishing the theoretical results, we introduce the following notation. 
We use $\mu_{\xh} := \e{f(\xh,\xit) - f(\xs,\xit)}$ to denote the optimality gap of $\xh$ and $\sigma^2_{\xh} := \var{f(\xh,\xit) - f(\xs,\xit)}$ to denote the associated variance term.
With i.i.d.\ sampling, $\e{\db_{lj}}=\e{\dbb_l}=\mu_{\xh}, \ \forall l,j$, $\var{\dbb_l}=\sigma^2_{\xh}/n_l, \ \forall l$, and under appropriate conditions, $\frac{n_l}{m_l} \var{n_l VD_l} \rightarrow \beta_{\gammab} \sigma^4_{\xh}$ as $l \rightarrow \infty$  by (\ref{eq:varvarred}).

@@

\begin{lemma} \label{lem:gbb_prob}
	Under assumptions A1 and A3, $\p{\dbb_l - \gbb_l \neq 0} \rightarrow 0$ exponentially fast as $l \rightarrow \infty$.
\end{lemma}

\begin{proof} When all the sampling problems used in the calculation of $\gbb$ have optimal solutions that solve (SP), i.e., when $\xs_{j} = \xs$, for all $j=1,2,\ldots, \nbl$, then $\gbb_l = \dbb_l$. 
Therefore, probability of the complement of this event provides an upper bound on the probability that $\dbb_l - \gbb_l \neq 0$. 
Let $E^C$ denote the complement of event $E$. 
Using this upper bound and continuing on, we obtain
	\begin{align}
		\p{\dbb_l - \gbb_l \neq 0} & \leq \p{\left( \bigcap_{j=1}^{\nbl} \{\xs_{j} = \xs\} \right)^C} \notag \\
		& = \p{ \bigcup_{j=1}^{\nbl} \{\xs_{j} \neq \xs\}} \notag \\
% 		& \leq \sum_{j=1}^{\nbl} \p{ \xs_j \neq \xs } \notag \\
		& \leq \left( \nbl\right) \p{ \xs_{m_l} \neq \xs }. \notag
	\end{align}
	With a batching structure in which $m_l = \lfloor n_l^{r} \rfloor$ for some $0<r<1$,  $\gamma_l \geq 1$, and by A3, the desired result follows.
\end{proof}

\begin{lemma} \label{lem:gb_gbb_l4}
	Under assumptions A1--A3, both (i) $\e{(\db_{lj} - \gb_{lj})^4} \rightarrow 0$ and (ii) $\e{(\dbb_l - \gbb_l)^4} \rightarrow 0$ exponentially fast as $l \rightarrow \infty$.
\end{lemma}

\begin{proof}
	For brevity, we drop the $lj$ subscripts from $\db_{lj}$ and $\gb_{lj}$ and $l$ from $\gbb_l$ and $\dbb_l$.
	(i) We can decompose the expectation as
	\begin{align*}
		\e{(\db - \gb)^4} & = \e{ \left.(\db - \gb)^4 \right| \db - \gb \neq 0} \p{\db - \gb \neq 0} + 0\\
		& \leq \e{ \left.(\db - \gb)^4 \right| \db - \gb \neq 0} \p{ \xs_{m_l}  \neq \xs}.
	\end{align*}
	The expectation term on the right-hand side is bounded by assumption A2 for i.i.d.\ samples.  
        Assumption A3 states $\p{\xs_{m_l} \neq \xs} \rightarrow 0$ exponentially fast.  
        Thus $\e{(\db - \gb)^4} \rightarrow 0$ exponentially fast.
(ii) This expectation term can be decomposed in the same manner
	\begin{align*}
		\e{(\dbb - \gbb)^4} & = \e{\left. (\dbb - \gbb)^4 \right| \dbb - \gbb \neq 0} \p{\dbb - \gbb \neq 0}.
	\end{align*}
	With the expectation term on the right-hand side again bounded by A2, and the probability term converges to zero exponentially fast by Lemma \ref{lem:gbb_prob}.
\end{proof}

@@

\begin{theorem} \label{thm:strong_consistency}
	Under assumptions A1--A3, (i) $\gbb_l - \dbb_l \rightarrow 0$, a.s., and (ii) $n_l VG_l - n_l VD_l \rightarrow 0$, a.s.\ as $l \rightarrow \infty$.  
        Therefore, $\gbb_l$ is a consistent estimator of $\mu_{\xh}$, and if $n_l VD_l$ is a consistent estimator of $\sigma^2_{\xh}$, then $n_l VG_l$ is a consistent estimator of $\sigma^2_{\xh}$.
\end{theorem}

\begin{proof}
	(i) By Lemma \ref{lem:gbb_prob}, $\p{\dbb_l - \gbb_l \neq 0}$ $\rightarrow 0$ exponentially fast.  
        Thus $\sum_l \p{\dbb_l - \gbb_l \neq 0} < \infty$.  
        The desired result follows by the Borel-Cantelli Lemma.
%
(ii) This proof is nearly identical to above. 
Similar to the proof of Lemma \ref{lem:gbb_prob}, we have
	\[
		\p{n_l VD_l -n_l VG_l \neq 0} \leq \p{\left( \bigcap_{j=1}^{\nbl} \{\xs_{j} = \xs\} \right)^C}.
	\]
	Following the same logic, $\p{n_l VD_l -n_l VG_l \neq 0} \rightarrow 0$ exponentially fast. 
Thus, $n_l VG_l$ is a consistent estimator of $\sigma^2_{\xh}$, if $n_l VD_l$ is, by the Borel-Cantelli Lemma.
\end{proof}

Part (i) of Theorem \ref{thm:strong_consistency} provides conditions under which the point estimator $\gbb$ of OMRP is a consistent estimator of the optimality gap, by establishing almost sure convergence of $\gbb_l - \dbb_l$ to $0$ as $l \rightarrow \infty$. 
Notice that under assumption A1, $\dbb_l \rightarrow \mu_\xh$, a.s. by the strong law of large numbers; as a result, $\gbb \rightarrow \mu_\xh$, a.s. 
Similarly, part (ii) of Theorem \ref{thm:strong_consistency} establishes almost sure convergence of $n_l VG_l - n_l VD_l$ to $0$ as $l \rightarrow \infty$. 
As a result, if $n_l VD_l$ is a consistent estimator of $\sigma^2_{\xh}$, so is $n_l VG_l$.  
Strong consistency of the variance estimator $n_l VD_l$ has been studied in \citep{damerdji1994strong} for the cases of nonoverlapping and maximally overlapping batches.  
Under the additional moment assumption
\begin{description}
	\item[A2$^\prime$]  $\exists \epsilon > 0$ such that $\sup_{\x \in X} \e{f(\x,\xi)^{4+\epsilon}} < \infty$,
\end{description}
for a sampling scheme of $m_l = \lfloor n_l^r \rfloor$, Corollaries 3.4 and 4.2 of \citep{damerdji1994strong} show that $n_l VD_l$ is consistent if $\tfrac{2}{4+\epsilon} < r < 1$ for nonoverlapping batches and $\tfrac{2}{4+\epsilon} < r < \tfrac{1}{2}$ for maximally overlapping batches.

@@

\begin{theorem} \label{thm:varvar_conv}
	Suppose assumptions A1--A3 hold, $\gammab = \frac{\gamma_l}{m_l}$ is constant for all $l$, and the batching structure ensures  $\frac{n_l}{m_l}\var{n_l VD_l} \rightarrow \beta_{\gammab} \sigma^4_{\xh}$ where $\beta_{\gammab}$ is the twice the value of (\ref{eq:var_reduct_formula}). 
        Then, $\frac{n_l}{m_l}\var{n_l VG_l} \rightarrow \beta_{\gammab} \sigma^4_{\xh}$ as $l \rightarrow \infty$ with the same $\beta_{\gammab}$. 
\end{theorem}

@@

	The fringe batches come in identically distributed pairs.  
        The first and last batch are identically distributed, as are the second and second-to-last, etc.  
        We know this because we could reverse the order of the i.i.d.\ sample and perform the batching again to get the exact same batch structure.  
		  The center batches are identically distributed.  
        Let $A_j = (\gb_j - \gbb)^2 - (\db_j - \dbb)^2$.
		  Furthermore, let $A_j^c$, $j=\left\lceil \gammab^{-1} \right\rceil+1, \ldots, n_b - \left\lceil \gammab^{-1} \right\rceil$ denote the center batches, and let $A_j^f$, $j=1,\ldots,\left\lceil \gammab^{-1} \right\rceil$ and $A_j^e$, $j=n_b - \left\lceil \gammab^{-1} \right\rceil+1,\ldots,n_b$ denote the fringe batches at the front and end of each set of nonoverlapping batches, respectively.  
		  Continuing from (\ref{eq:vvar_conv_pf_1}),
	\begin{align*}
		& \left( \frac{n}{dt} \right)^2 n_b \e{ \sum_{j=1}^{n_b} \left[ (\gb_j - \gbb)^2 - (\db_j - \dbb)^2 \right]^2 } \\
		 =  & \left( \frac{n}{dt} \right)^2 n_b \e{\sum_{j=1}^{\left\lceil \gammab^{-1} \right\rceil} \left( A_j^f \right)^2 \
		 + \sum_{j=\left\lceil \gammab^{-1} \right\rceil+1}^{n_b-\left\lceil \gammab^{-1} \right\rceil} \left( A_j^c \right)^2 \
		 + \sum_{j=n_b-\left\lceil \gammab^{-1} \right\rceil+1}^{n_b} \left( A_j^e \right)^2 
		} \\
%		 = & \left( \frac{n}{dt} \right)^2 n_b \sum_{j=\left\lceil \gammab^{-1} \right\rceil+1}^{n_b-\left\lceil \gammab^{-1} \right\rceil} \e{ \left( A_j^c \right)^2 } + \left( \frac{n}{dt} \right)^2 (n_b) 2\sum_{j=1}^{\left\lceil \gammab^{-1} \right\rceil} \e{ \left( A_j^f \right)^2 } \\
		 = & \left( \frac{n}{dt} \right)^2 n_b \left(n_b - 2\left\lceil \gammab^{-1} \right\rceil \right) \e{ \left( A_j^c \right)^2 }
		+ \left( \frac{n}{dt} \right)^2 (n_b) 2\sum_{j=1}^{\left\lceil \gammab^{-1} \right\rceil} \e{ \left( A_j^f \right)^2 }.
	\end{align*}	
	The last equality follows because each front-fringe batch's $\e{(A_j^f)^2}$, $j=1,\ldots,\left\lceil \gammab^{-1} \right\rceil$  has an identical pair in an end-fringe batch's $\e{(A_j^e)^2}$, $j=n_b - \left\lceil \gammab^{-1} \right\rceil+1,\ldots,n_b$, and all center batches have identical values of $\e{(A_j^c)^2}$, $j=\left\lceil \gammab^{-1} \right\rceil+1, \ldots, n_b - \left\lceil \gammab^{-1} \right\rceil$.  
	As a result, we have $\left\lceil \gammab^{-1} \right\rceil + 1$ expectation terms, each of which has a coefficient that is less than $\left(\frac{n}{dt}\right)^2n_b^2 < (2m)^2$.  
        Removing the notation for fringe or center batch for simplicity, each expectation can be bounded by
	\begin{align}
		\left( 2m \right)^2 \e{ \left( A_j \right)^2 } & = \left( 2m \right)^2 \e{ \left((\gb_j - \gbb) - (\db_j - \dbb)\right)^2 \left((\gb_j - \gbb) + (\db_j - \dbb)\right)^2 } \notag \\
		& \leq \left( 2m \right)^2 \sqrt{\e{\left((\gb_j - \gbb) - (\db_j - \dbb)\right)^4}} \sqrt{\e{\left((\gb_j - \gbb) + (\db_j - \dbb)\right)^4 }} \label{eq:var_reduct_cauchy2} \\
		& \leq \left( 2m \right)^2 \sqrt{\e{\left((\gb_j - \gbb) - (\db_j - \dbb)\right)^4}} \sqrt{8 \left(\e{(\gb_j - \gbb)^4} + \e{(\db_j - \dbb)^4} \right)}, \label{eq:var_reduct_holder2}
	\end{align}
	where equation (\ref{eq:var_reduct_cauchy2}) follows from the Cauchy-Schwartz inequality, and equation (\ref{eq:var_reduct_holder2}) follows from H\"{o}lder's inequality.  
        Since we can bound both expectation terms in (\ref{eq:var_reduct_holder2}) by A2, we can demonstrate $\sqrt{\frac{n}{m}}(nVG - nVD) \xrightarrow{L^2} 0$ by focusing on only one of the terms.  
        Focusing on the first expectation term in (\ref{eq:var_reduct_holder2}) and returning the $\tfrac{n}{m}$, by H\"{o}lder's inequality,
	\begin{align*}

>>>>

In order to apply overlapping batches to MRP, we need to keep track of solutions to sampling problems (SP$_m$) for each batch of size $m$.  
Toward this end, let $B(i)$ denote the set of batches $j \in \left\{1, 2, \dots, \nb \right\}$ observation $\xiti$ is used in, $i = 1, 2, \dots, n$.  
See Figure \ref{fig:overlap_nonint} for an example with $n=18$, $m = 6$, and $\gamma = 2$.  
Here, the first batch uses observations $\xit^1, \xit^2, \xit^3, \xit^4, \xit^5, \xit^6$, the second batch uses $\xit^3, \xit^4, \xit^5, \xit^6, \xit^7, \xit^8$, and so on.  
So, $B(1) = \{1\}$, $B(3) = \{1,2\}$, and $B(7)=\{2,3,4\}$.  
As before, we use $\xs_j$ to denote an optimal solution to sampling problem (SP$_m$) formed using the $j$th batch, which may not be unique; so we set $\xs_j \in \argmin_{\x \in X} \frac{1}{m} \sum_{i=1}^m f(\x,\xit^{\gamma(j-1) + i})$, $j = 1, 2, \dots, \nb$.

@@

The optimality gap estimator for each batch $\gb_{lj}$ is defined like (\ref{eq:mrp_gap}) for general values of the nonoverlap parameter $\gamma_l$.  
We simply removed the minimization in (\ref{eq:mrp_gap}) and directly used an optimal solution of $\xs_j$ of batch $j$.  
The overall mean, $\gbb_l$, is defined a little differently.  
Here, $\gbb_l$ still uses each observation $i = 1, 2, \dots, n_l$ but also makes use of {\it all} the information collected throughout the batches.  
That is, if observation $\xiti$ is used in $|B(i)|$ number of batches, all the optimal solutions $\xs_j$ corresponding to each batch $j \in B(i)$ are used for the lower bound estimator.  
Then, $VG_l$ is defined in a similar fashion for the overlapping batches variance estimator (\ref{eq:variance_gamma}).

The motivation behind forming $\gbb_l$ by taking into account $B(i)$ is to give each observation $\xit^i$ equal weight. 
This is because we want to be as close to a standard mean estimator (e.g., $\ybb=\frac{1}{n}\sum_{i=1}^{n} y^i$) as possible.
With this definition of $\gbb_l$, when $\xs_j= \xs$ for all $j= 1, 2, \dots, \nbl$, we have $\gbb_l$ equal to its nonoptimized counterpart given in (\ref{eq:dbb}), which is the standard mean estimator. 
This fact is key in establishing the theoretical results presented in the next section. 
This choice of $\gbb_l$, however, may not be guaranteed to be nonnegative.
In practice, $\max\{\gbb_l,0\}$ can be used as the point estimator of the optimality gap to ensure nonnegativity, and $VG_l$ can be calculated the same way given in (\ref{eq:vg}), using $\gbb_l$, to obtain improved interval estimators via overlapping.

@@

In this section, we provide conditions under which several asymptotic properties hold for OMRP estimators. 
First, we show the consistency of the point estimators of the optimality gap, $\gbb_l$, and variance, $n_l VG_l$. 
Then, we provide conditions under which the OMRP variance estimator achieves asymptotically lower variances relative to the MRP estimator.  
Finally, we show the asymptotic validity of the OMRP interval estimator.

@@

\begin{description}
	\item[A1] Samples of the random variable $\xit$ are i.i.d.
	\item[A2] $\sup_{\x \in X} \e{f(\x,\xit)^{4}} < \infty$.
	\item[A3] (\ref{eq:sto_prog}) has a unique optimal solution $\xs$ and $\p{\xs_m \neq \xs} \rightarrow 0$ exponentially fast as $m \rightarrow \infty$.  
           That is, there exist constants $M > 0$ and $c > 0$ such that $\log(\p{\xs_m \neq \xs}) \leq -cm$ for all $m > M$.
\end{description}

First, we restrict our attention to i.i.d.\ sampling by A1.
The results for overlapping batches from simulation output analysis indicate that $\var{n_l VG_l}$ may be reduced by overlapping, thus requiring at least finite fourth moments. 
We will formally show this result in \S\ref{ssec:var_reduct}.
Our proof relies on an exponential rate of convergence of the sampled solution to the true solution of the problem, specified by A3.  
Assumption A3 can be satisfied for different classes of problems with unique optimal solutions.  
\citet{shapiro2000rate} establish exponential rates of convergence for a class of (\ref{eq:sto_prog}), which include two-stage stochastic linear programs with recourse when the support of $\xit$ is finite.   
\citet{kleywegt2002sample} show that Assumption A3 can be satisfied for discrete stochastic programs (i.e., when $X$ is a finite set) under the condition that the moment generating function of $f(\xs,\xit) - f(\x,\xit)$ for $\x \in X \backslash \{\xs\}$ is finite valued on $\R$.  
This class of problems includes stochastic integer programs with recourse.
  We present computational results on both stochastic linear and integer programs with recourse in \S \ref{sec:comp}.

@@

Before we begin establishing the theoretical results, we introduce the following notation. 
We use $\mu_{\xh} = \e{f(\xh,\xit) - f(\xs,\xit)}$ to denote the optimality gap of $\xh$ and $\sigma^2_{\xh} = \var{f(\xh,\xit) - f(\xs,\xit)}$ to denote the associated variance term.
With i.i.d.\ sampling, $\e{\db_{lj}}=\e{\dbb_l}=\mu_{\xh}, \ \forall l,j$, $\mathrm{Var}(\dbb_l)=\sigma^2_{\xh}/n_l, \ \forall l$, and under appropriate conditions, $\frac{n_l}{m_l} \var{n_l VD_l} \rightarrow \beta_{\gammab} \sigma^4_{\xh}$ as $l \rightarrow \infty$  by (\ref{eq:varvarred}).

@@

\begin{lemma} \label{lem:gbb_prob}
	Under Assumptions A1 and A3, $\p{\dbb_l - \gbb_l \neq 0} \rightarrow 0$ exponentially fast as $l \rightarrow \infty$.
\end{lemma}

\begin{proof} When all the sampling problems used in the calculation of $\gbb$ have optimal solutions that solve (SP), i.e., when $\xs_{j} = \xs$, for all $j=1,2,\ldots, \nbl$, then $\gbb_l = \dbb_l$. 
Therefore, the probability of the complement of this event provides an upper bound on the probability that $\dbb_l - \gbb_l \neq 0$. 
Let $E^C$ denote the complement of event $E$. 
Using this upper bound and continuing on, we obtain
	\begin{align}
		\p{\dbb_l - \gbb_l \neq 0} & \leq \p{\left( \bigcap_{j=1}^{\nbl} \{\xs_{j} = \xs\} \right)^C} \notag \\
		& = \p{ \bigcup_{j=1}^{\nbl} \{\xs_{j} \neq \xs\}} \notag \\
		& \leq \left( \nbl\right) \p{ \xs_{m_l} \neq \xs }. \notag
	\end{align}
	With a batching structure in which $m_l = \lfloor n_l^{r} \rfloor$ for some $0<r<1$,  $\gamma_l \geq 1$, and by A3, the desired result follows.
\end{proof}

\begin{lemma} \label{lem:gb_gbb_l4}
	Under Assumptions A1--A3, both (i) $\e{(\db_{lj} - \gb_{lj})^4} \rightarrow 0$ and (ii) $\e{(\dbb_l - \gbb_l)^4} \rightarrow 0$ exponentially fast as $l \rightarrow \infty$.
\end{lemma}

\begin{proof}
	For brevity, we drop the $lj$ subscripts from $\db_{lj}$ and $\gb_{lj}$ and $l$ from $\gbb_l$ and $\dbb_l$.
	(i) We can decompose the expectation as
	\begin{align*}
		\e{(\db - \gb)^4} & = \e{ \left.(\db - \gb)^4 \right| \db - \gb \neq 0} \p{\db - \gb \neq 0} + 0\\
		& \leq \e{ \left.(\db - \gb)^4 \right| \db - \gb \neq 0} \p{ \xs_{m_l}  \neq \xs}.
	\end{align*}
	The expectation term on the right-hand side is bounded by Assumption A2 for i.i.d.\ samples.  
        Assumption A3 states $\p{\xs_{m_l} \neq \xs} \rightarrow 0$ exponentially fast.  
        Thus $\e{(\db - \gb)^4} \rightarrow 0$ exponentially fast.
(ii) This expectation term can be decomposed in the same manner as
$
		\e{(\dbb - \gbb)^4}  = \e{\left. (\dbb - \gbb)^4 \right| \dbb - \gbb \neq 0} \p{\dbb - \gbb \neq 0}.
$ 
	The proof follows using similar arguments as proof of part (i) and Lemma \ref{lem:gbb_prob}.
\end{proof}

@@

\begin{theorem} \label{thm:strong_consistency}
	Under Assumptions A1--A3, (i) $\gbb_l - \dbb_l \rightarrow 0$, a.s., and (ii) $n_l VG_l - n_l VD_l \rightarrow 0$, a.s.\ as $l \rightarrow \infty$.  
        Therefore, $\gbb_l$ is a consistent estimator of $\mu_{\xh}$, and if $n_l VD_l$ is a consistent estimator of $\sigma^2_{\xh}$, then $n_l VG_l$ is a consistent estimator of $\sigma^2_{\xh}$.
\end{theorem}

\begin{proof}
	(i) By Lemma \ref{lem:gbb_prob}, $\p{\dbb_l - \gbb_l \neq 0}$ $\rightarrow 0$ exponentially fast.  
        Thus $\sum_l \p{\dbb_l - \gbb_l \neq 0} < \infty$.  
        The desired result follows by the Borel-Cantelli Lemma.
%
(ii) Similar to the proof of Lemma \ref{lem:gbb_prob}, we have
	\[
		\p{n_l VD_l -n_l VG_l \neq 0} \leq \p{\left( \bigcap_{j=1}^{\nbl} \{\xs_{j} = \xs\} \right)^C}.
	\]
	Following the same logic as in (i), $\p{n_l VD_l -n_l VG_l \neq 0} \rightarrow 0$ exponentially fast. 
Thus, $n_l VG_l$ is a consistent estimator of $\sigma^2_{\xh}$, if $n_l VD_l$ is, by the Borel-Cantelli Lemma.
\end{proof}

Part (i) of Theorem \ref{thm:strong_consistency} provides conditions under which the point estimator $\gbb$ of OMRP is a consistent estimator of the optimality gap, by establishing almost sure convergence of $\gbb_l - \dbb_l$ to $0$ as $l \rightarrow \infty$. 
Notice that under Assumption A1, $\dbb_l \rightarrow \mu_\xh$, a.s. by the strong law of large numbers; as a result, $\gbb \rightarrow \mu_\xh$, a.s. 
Similarly, part (ii) of Theorem \ref{thm:strong_consistency} establishes almost sure convergence of $n_l VG_l - n_l VD_l$ to $0$ as $l \rightarrow \infty$. 
As a result, if $n_l VD_l$ is a consistent estimator of $\sigma^2_{\xh}$, so is $n_l VG_l$.  
Strong consistency of the variance estimator $n_l VD_l$ has been studied in \cite{damerdji1994strong} for the cases of nonoverlapping and maximally overlapping batches.  
Under the additional moment assumption
\begin{description}
	\item[A2$^\prime$]  $\exists \epsilon > 0$ such that $\sup_{\x \in X} \e{f(\x,\xi)^{4+\epsilon}} < \infty$,
\end{description}
for a sampling scheme of $m_l = \lfloor n_l^r \rfloor$, Corollaries 3.4 and 4.2 of \cite{damerdji1994strong} show that $n_l VD_l$ is consistent if $\tfrac{2}{4+\epsilon} < r < 1$ for nonoverlapping batches and $\tfrac{2}{4+\epsilon} < r < \tfrac{1}{2}$ for maximally overlapping batches.

@@

\begin{theorem} \label{thm:varvar_conv}
	Suppose Assumptions A1--A3 hold, $\gammab = \frac{\gamma_l}{m_l}$ is constant for all $l$, and the batching structure ensures  $\frac{n_l}{m_l}\var{n_l VD_l} \rightarrow \beta_{\gammab} \sigma^4_{\xh}$ where $\beta_{\gammab}$ is the twice the value of (\ref{eq:var_reduct_formula}). 
        Then, $\frac{n_l}{m_l}\var{n_l VG_l} \rightarrow \beta_{\gammab} \sigma^4_{\xh}$ as $l \rightarrow \infty$ with the same $\beta_{\gammab}$. 
\end{theorem}

@@

	The fringe batches come in identically distributed pairs.  
        The first and last batch are identically distributed, as are the second and second-to-last, etc.  
        We know this because we could reverse the order of the i.i.d.\ sample and perform the batching again to get the same batch structure.  
		  The center batches are identically distributed.  
        Let $A_j = (\gb_j - \gbb)^2 - (\db_j - \dbb)^2$.
		  Furthermore, let $A_j^c$, $j=\left\lceil \gammab^{-1} \right\rceil+1, \ldots, n_b - \left\lceil \gammab^{-1} \right\rceil$ denote the center batches, and let $A_j^f$, $j=1,\ldots,\left\lceil \gammab^{-1} \right\rceil$ and $A_j^e$, $j=n_b - \left\lceil \gammab^{-1} \right\rceil+1,\ldots,n_b$ denote the fringe batches at the front and end of each set of nonoverlapping batches, respectively.  
		  Continuing from (\ref{eq:vvar_conv_pf_1}),
	\begin{align*}
		& \left( \frac{n}{dt} \right)^2 n_b \e{ \sum_{j=1}^{n_b} \left[ (\gb_j - \gbb)^2 - (\db_j - \dbb)^2 \right]^2 } \\
		 =  & \left( \frac{n}{dt} \right)^2 n_b \e{\sum_{j=1}^{\left\lceil \gammab^{-1} \right\rceil} \left( A_j^f \right)^2 \
		 + \sum_{j=\left\lceil \gammab^{-1} \right\rceil+1}^{n_b-\left\lceil \gammab^{-1} \right\rceil} \left( A_j^c \right)^2 \
		 + \sum_{j=n_b-\left\lceil \gammab^{-1} \right\rceil+1}^{n_b} \left( A_j^e \right)^2 
		} \\
		 = & \left( \frac{n}{dt} \right)^2 n_b \left(n_b - 2\left\lceil \gammab^{-1} \right\rceil \right) \e{ \left( A_j^c \right)^2 }
		+ \left( \frac{n}{dt} \right)^2  2 n_b \sum_{j=1}^{\left\lceil \gammab^{-1} \right\rceil} \e{ \left( A_j^f \right)^2 }.
	\end{align*}	
	The last equality follows because each front-fringe batch's $\e{(A_j^f)^2}$, $j=1,\ldots,\left\lceil \gammab^{-1} \right\rceil$  has an identical pair in an end-fringe batch's $\e{(A_j^e)^2}$, $j=n_b - \left\lceil \gammab^{-1} \right\rceil+1,\ldots,n_b$, and all center batches have identical values of $\e{(A_j^c)^2}$, $j=\left\lceil \gammab^{-1} \right\rceil+1, \ldots, n_b - \left\lceil \gammab^{-1} \right\rceil$.  
	As a result, we have $\left\lceil \gammab^{-1} \right\rceil + 1$ expectation terms, each of which has a coefficient that is less than $\left(\frac{n}{dt}\right)^2n_b^2 < (2m)^2$.  
        Removing the notation for fringe or center batch for simplicity, each expectation can be bounded by
	\begin{align}
		\left( 2m \right)^2 \e{ \left( A_j \right)^2 } & = \left( 2m \right)^2 \e{ \left((\gb_j - \gbb) - (\db_j - \dbb)\right)^2 \left((\gb_j - \gbb) + (\db_j - \dbb)\right)^2 } \notag \\
		& \leq \left( 2m \right)^2 \sqrt{\e{\left((\gb_j - \gbb) - (\db_j - \dbb)\right)^4}} \sqrt{\e{\left((\gb_j - \gbb) + (\db_j - \dbb)\right)^4 }} \label{eq:var_reduct_cauchy2} \\
		& \leq \left( 2m \right)^2 \sqrt{\e{\left((\gb_j - \gbb) - (\db_j - \dbb)\right)^4}} \sqrt{8 \left(\e{(\gb_j - \gbb)^4} + \e{(\db_j - \dbb)^4} \right)}, \label{eq:var_reduct_holder2}
	\end{align}
	where (\ref{eq:var_reduct_cauchy2}) follows from the Cauchy-Schwartz inequality, and (\ref{eq:var_reduct_holder2}) follows from H\"{o}lder's inequality.  
        Since we can bound both expectation terms in (\ref{eq:var_reduct_holder2}) by A2, we can demonstrate $\sqrt{\frac{n}{m}}(nVG - nVD) \xrightarrow{L^2} 0$ by focusing on only one of the terms.  
        Focusing on the first expectation term in (\ref{eq:var_reduct_holder2}) and returning the $\tfrac{n}{m}$, by H\"{o}lder's inequality,
	\begin{align*}
		\left(\frac{n}{m} 4m^2\right)^2 \e{\left((\gb_j - \gbb) - (\db_j - \dbb)\right)^4}	& \leq 8 \left( \frac{n}{m} 4m^2\right)^2 \left( \e{(\db_j - \gb_j)^4} + \e{(\dbb - \gbb)^4} \right).
	\end{align*}	
	By Lemma \ref{lem:gb_gbb_l4}, both expectation terms converge to zero with exponential rates.  
        With a batching structure $m_l = \lfloor n_l^r \rfloor$ for some $0<r<1$, we see that $\frac{n_l}{m_l} \e{(n_l VG_l - n_l VD_l)^2} \rightarrow 0$.
\end{proof}
